Shades of Green in the Swedish Forest-Based Bioeconomy Transition

### 2023-09-22 docs/reports/paper/paper1.qmd
update figure2 bv count

---
title: The Direction of Search for a Forest Based Bioeconomy
author: 
    name: Jonas Kreutzer
    affiliation: Department of Economic History, Lund University
date: last-modified
abstract: |
    The transition to a bioeconomy is high on the agenda for research, industry and policy. Beyond conceptual uncertainties of what is meant by a bioeconomy, three different visions have emerged, differing in their aims and objectives, drivers and mediators of innovation and relation to the biosphere. The use of detailed innovation data from a Swedish innovation database, SWINNO, allows to capture the actual innovation output of the Swedish forest based bioeconomy innovation system. Qualitative coding of the underlying source material associates innovation from 1970 to 2020 with vision of a bioresource, biotechnology and / or bioecology bioeconomy. Preliminary results suggest that the innovation output of the forest based bioeconomy has failed to gain in prominence. Of the three visions, the bioresource vision has been the most dominant, especially in the 1970s and after 2005. At the same time, innovation which fit into multiple visions exist, suggesting that the opportunity space for future development is not yet locked in.

@@

Innovations were classified as belonging to a bioeconomy vision if they fulfilled one, or more criteria outlined in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy.
@tbl-bugge and @tbl-vivien show the summary of bioeconomy visions presented in the cited papers.
For the remainder of the paper the nomenclature of @bugge2016WhatBioeconomyReview is adopted.
Vivien et al.'s [-@vivien2019HijackingBioeconomy] Type I bioeconomy is taken to correspond with the bioecology vision, Type II and Type III to correspond with the bioresource and biotechnology vision, respectively.
The analysis of bioeconomy discourse which revealed different visions, highlights that the visions have emerged at different times and with different actors at their centers [@vivien2019HijackingBioeconomy, @bugge2016WhatBioeconomyReview].
@vivien2019HijackingBioeconomy trace the introduction of the term bioeconomics to the 1920s and a work on fishery by the Russian biologist Baranoff.
The term bioeoconomics which they label as Type I bioeconomy, was introduced by Georgescu-Roegen,  in -@roegen1975BioeconomicAspectsEntropy.
@vivien2019HijackingBioeconomy argue that the term biotechnology gradually emerged between 1990 and 2000, in the wake of enthusiasm about biotechnological adavances, such as the discovery of DNA, in preceeding years.
It gained in popularity and international importance when the OECD began increasingly supporting and reshaping its meaning at the turn of the millenium [@vivien2019HijackingBioeconomy].
According to them the last concept to enter the discourse was the bioresource oriented bioeconomy.
While biomass has been used throughout history in various ways, the renewed interest with the intention to replace fossil inputs, for example in European Comission communication since 2010.
However, @vivien2019HijackingBioeconomy point to the long history of biorefinery as a core aspect of this vision, which was discussed as early as the 1930s.
Hence, while the organization of the discourse into the labels biotechnology, bioecology and bioresource vision is a very recent research outcome, the underlying concepts have a richer history.
This alleviates concerns that using these labels retroactively confines the innovations to current policy terms.

@@

```{python}
# | label: fig-bioeconomy_trends
# | fig-cap: Bioeconomy Vision Trends

@@

# Discussion

>>>>

---
title: Shades of Green in the Swedish Forest-Based Bioeconomy Transition 
subtitle: Draft
author: 
    name: Jonas Kreutzer
    affiliation: Department of Economic History, Lund University
date: last-modified
abstract: |
    The transition to a bioeconomy is high on the agenda for research, industry and policy. Beyond conceptual uncertainties of what is meant by a bioeconomy, three different visions have emerged, differing in their aims and objectives, drivers and mediators of innovation and relation to the biosphere. The use of detailed innovation data from a Swedish innovation database, SWINNO, allows to capture the actual innovation output of the Swedish forest based bioeconomy innovation system. Qualitative coding of the underlying source material associates innovation from 1970 to 2020 with vision of a bioresource, biotechnology and / or bioecology bioeconomy. Preliminary results suggest that the innovation output of the forest based bioeconomy has failed to gain in prominence. Of the three visions, the bioresource vision has been the most dominant, especially in the 1970s and after 2005. At the same time, innovation which fit into multiple visions exist, suggesting that the opportunity space for future development is not yet locked in.

@@

Innovations were classified as belonging to a bioeconomy vision if they fulfilled one, or more criteria outlined in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy.
@tbl-bugge and @tbl-vivien show the summary of bioeconomy visions presented in the cited papers.
For the remainder of the paper the nomenclature of @bugge2016WhatBioeconomyReview is adopted.
Vivien et al.'s [-@vivien2019HijackingBioeconomy] Type I bioeconomy is taken to correspond with the bioecology vision, Type II and Type III to correspond with the bioresource and biotechnology vision, respectively.
The analysis of bioeconomy discourse which revealed different visions, highlights that the visions have emerged at different times and with different actors at their centers [@vivien2019HijackingBioeconomy, @bugge2016WhatBioeconomyReview].
@vivien2019HijackingBioeconomy trace the introduction of the term bioeconomics to the 1920s and a work on fishery by the Russian biologist Baranoff.
The term bioeoconomics which they label as Type I bioeconomy, was introduced by Georgescu-Roegen,  in -@georgescu-roegen1975BioeconomicAspectsEntropy.
@vivien2019HijackingBioeconomy argue that the term biotechnology gradually emerged between 1990 and 2000, in the wake of enthusiasm about biotechnological adavances, such as the discovery of DNA, in preceeding years.
It gained in popularity and international importance when the OECD began increasingly supporting and reshaping its meaning at the turn of the millenium [@vivien2019HijackingBioeconomy].
According to them the last concept to enter the discourse was the bioresource oriented bioeconomy.
While biomass has been used throughout history in various ways, the renewed interest with the intention to replace fossil inputs, for example in European Comission communication since 2010.
However, @vivien2019HijackingBioeconomy point to the long history of biorefinery as a core aspect of this vision, which was discussed as early as the 1930s.
Hence, while the organization of the discourse into the labels biotechnology, bioecology and bioresource vision is a very recent research outcome, the underlying concepts have a richer history.
This alleviates concerns that using these labels retroactively confines the innovations to current policy terms.

@@

```{python}
# | label: fig-bioeconomy_trends
# | fig-cap: "Count of Bioeconomy Visions by Year"

@@
